Think deep for reef resilience

### 2025-07-18 paper/paper_draft.qmd
Add section on connectivity improvements for further analyses
Currently unclear of what 'improvements' refers to other than larger datasets

Five different GCMs were included in analyses to capture uncertainty in climate system
responses to increasing atmospheric carbon dioxide levels. The large differences in cover
timeseries seen across the GCMs suggest that it is important to consider this uncertainty in
analyses of biological climate change responses. This is supported by recent literature
suggesting an effective approach to capturing climate change uncertainty is to model climate
change responses using a range of climate models separately [@shepherd2018]. However, the
patterns identified between low, medium and high clusters of reefs suggest that the
importance of depth refugia is consistent across scenarios. Similarly, at the management
area scale, only 1% of reefs were assigned a ‘low’ and ‘high’ cluster in different GCMs,
suggesting that intrinsic reef properties can prevent high cover resilient reefs from
crashing under a different future climate scenario.

There are multiple caveats in how coral ecosystem dynamics are modelled in this
investigation. We model coral colonies as objects with circular diameter values and track
these diameters over timesteps through growth. Due to the availability of quality future
projected data we do not consider the impacts of waves, cyclones and currents on coral
colony survival and bleaching (insert citations for waves affecting survival and bleaching).
We do not consider the impacts of Crown of Thorns Starfish (CoTS) on coral survival and reef
resilience (insert citation for cots affecting corals). Furthermore, there are multiple
caveats to acknowledge in the data sources used in this investigation. Uncertainty in
connectivity modelling outputs, satellite-derived bathymetry data and future thermal stress
projections are likely to influence the results obtained in this investigation. To limit the
influence of bathymetry data on results, reefs with less than 5% data coverage are removed
from analyses. The DHW data used in this investigation have not been bias corrected for the
Great Barrier Reef region and is part of a wider Australian dataset [@dixon2022] developed
using statistical downscaling. We investigate a range of future climate responses and
thermal projections to account for a range of possible futures, and our results suggest
the factors that influence reef resilience are consistent across these scenarios.

>>>>

Five different GCMs were included in analyses to capture uncertainty in climate system
responses to increasing atmospheric carbon dioxide levels. The large differences in cover
timeseries seen across the GCMs suggest that it is important to consider this uncertainty in
analyses of biological climate change responses. This is supported by recent literature
suggesting an effective approach to capturing climate change uncertainty is to model climate
change responses using a range of climate models separately [@shepherd2018]. However, the
patterns identified between low, medium and high clusters of reefs suggest that the
importance of depth refugia is consistent across scenarios. 
<!-- Similarly, at the management
area scale, only 1% of reefs were assigned a ‘low’ and ‘high’ cluster in different GCMs,
suggesting that intrinsic reef properties can prevent high cover resilient reefs from
crashing under a different future climate scenario. -->

There are multiple caveats in how coral ecosystem dynamics are modelled in this
investigation. We model coral colonies as objects with circular diameter values and track
these diameters over timesteps through growth. Due to the availability of quality future
projected data we do not consider the impacts of waves, cyclones and currents on coral
colony survival and bleaching (insert citations for waves affecting survival and bleaching).
We do not consider the impacts of Crown of Thorns Starfish (CoTS) on coral survival and reef
resilience (insert citation for cots affecting corals). Furthermore, there are multiple
caveats to acknowledge in the data sources used in this investigation. Uncertainty in
connectivity modelling outputs, satellite-derived bathymetry data and future thermal stress
projections are likely to influence the results obtained in this investigation. Incorporating 
further improvements in connectivity modelling and understanding of the influence of 
climate change on reef connectivity would improve certainty in the conclusions suggested 
in this investigation. To limit the influence of bathymetry data on results, reefs with 
less than 5% data coverage are removed from analyses. The DHW data used in this 
investigation have not been bias corrected for the Great Barrier Reef region and is part of
a wider Australian dataset [@dixon2022] developed using statistical downscaling. We 
investigate a range of future climate responses and thermal projections to account for 
a range of possible futures, and our results suggest the factors that influence reef 
resilience are consistent across these scenarios.
